Operationally-Safe Peer-to-Peer Energy Trading in Distribution Grids: A Game-Theoretic Market-Clearing Mechanism

### 2022-03-04 papers/02 journal v2/journal_v14.tex
g

Note that the cost functions in \eqref{eq:cost_def} are coupled only via the aggregative quantity $\sigma^{\textrm{mg}} = \sum_{i\in\mathcal{N}} p_{i}^{\mathrm{mg}}$ in  \eqref{eq:Aggr}, namely, the active load (i.e., the congestion) on the main grid. Therefore, for each agent $i \in \mc I$, we can define a function $\tilde J_i$ such that
\begin{equation}
\tilde J_i(u_i, \sigma^{\textrm{mg}}) =:J_i(u_i,  u_{-i}).
\end{equation}
Games with such special structure are known as \textit{aggregative games} \cite{jensen2010aggregative}, and have received intense research interest, within the operations research and the automatic control communities \cite{paccagnan2019,belgioioso2020semi,belgioioso2020b,
gadjov2020single,bianchi2020fast}. 
\marginnote{\note{R1-5}}
\edit{
When the agents' cost functions depend linearly on the congestion (as for our P2P market model), v-GNEs are efficient. Specifically, the so-called \textit{price of anarchy}\cite{koutsoupias1999worst}, which quantifies how much selfish behaviour degrades the performance of a given system, tends to one (i.e., no performance degradation) as the agents population size grows \cite{paccagnan2018efficiency}.}

@@

\item a (dual) variable $\mu^{\text{tr}}_{(i,j)}$ for each trading partner $j \in \mc N_i$, whose function is to drive prosumer $i$'s and $j$'s power trade to agreement, i.e., the reciprocity constraints \eqref{eq:rep_const}, and can be interpreted from an economic perspective as a bilateral trading shadow price \cite[Section 2.4]{lecadre2020}.

\end{itemize}
In addition to the physical variables of the distribution network, i.e., $u_{N\!+\!1}$, the DNO stores in its local memory
\begin{itemize}
\item the (dual) variables $\lambda^{\text{mg}}$ and $\mu^{\text{tg}} $, that are associated with the main grid constraints \eqref{eq:p_mg_bound} and \eqref{eq:grid_const2}, respectively;

\item a (dual) variable $\mu_{y}^{\text{pb}}$ for each bus $y \in \mc B$, associated with the power balance constraint on bus $y$ \eqref{eq:p_bal_p}.
\end{itemize}
From an economic perspective, these variables can be interpreted as extra marginal losses imposed to the prosumers for the grid usage. From a control-theoretic perspective, they can be interpreted as states of discrete-time integrators driven by the violation of the network operational constraints \eqref{eq:Cconstr_def}.
}
%
\begin{figure}[t!]
\centering
\includegraphics[width=.3\textwidth]{figures/37bus_sys_f7}
\smallskip
\caption{\edit{Information flow in Algorithm 1.}}
\label{fig:IFC}
\end{figure}%
%
\begin{figure}[t]
\marginnote{\note{R1-6}}
\edit{
\begin{minipage}{\columnwidth}
\hrule
\smallskip
\textsc{Algorithm 1}. Coordinated P2P Market Clearing
\smallskip
\hrule 
\smallskip
\text{Initialization}: For all prosumers $i \in \mc N$: set $\mu^{\text{tr}}_{(i,j)}(-1) = 0$, $\forall j \in \mc N_i$. DNO: set $\lambda^{\text{mg}}(0)\!=\!0$, $\mu^{\text{tg}}(0) \!=\! 0$, $\mu_y^{\text{pb}}(0)\! = \!0$, $\forall y \!\in\! \mc B$.

\smallskip
 \text{Iterate until convergence $(k = 0,1,\ldots)$}
\begin{align*}
& \text{For all prosumers $i \in \mc I$ (in parallel):}\\
\quad &\left\lfloor
\begin{array}{l}
\text{Local update via \textsc{Algorithm 2}:}\\
\left\lfloor
\begin{array}{l}
\text{Set $\{\mu_{(i,j)}(k)\}_{j \in \mc N_i}$ as in \textsc{Alg. 2} (i)}\\
\text{Set $u_i(k\!+\!1)$ as in \textsc{Alg. 2} (ii)}
\end{array}
\right.\\[1em]
 \text{Communication}\\
 \left\lfloor
\begin{array}{l}
\eta_i(k\!+\!1), \; p_i^{\text{mg}}(k\!+\!1) \rightarrow \text{ DNO}\\[.5em]
\text{For all trading partners } j \in \mc N_i \text{ (in parallel):}\\
 \left\lfloor
\begin{array}{l}
p_{(i,j)}^{\text{tr}}(k\!+\!1) \rightarrow \text{ prosumer } j
\end{array}
\right.
\end{array}
\right.
\vspace*{.1em}
\end{array}
\right.\\[.5em]
%%%
% DNO
%%%
& \text{DNO}\\
\quad &\left\lfloor
\begin{array}{l}
\text{Central update via \textsc{Algorithm 3}:}\\
\left\lfloor
\begin{array}{l}
\text{Set $u_{N+1}(k\!+\!1)$ as in \textsc{Alg. 3} (i)}\\
\text{Set $\lambda^{\text{mg}}$, $\mu^{\text{tg}}$, $\{ \mu_y^{\text{pb}}(k\!+\!1)\}_{y \in\mc B}$ as in \textsc{Alg. 3} (ii)}
\end{array}
\right.\\[1em]
\text{Communication (Gather \& Broadcast)}\\
 \left\lfloor
\begin{array}{l}
\sigma^{\mathrm{mg}},\, \lambda^{\text{mg}}, \mu^{\text{tg}}(k\!+\!1) \rightarrow \text{ all prosumers } i \in \mc N\\[.5em]
\text{For all busses } y \in \mc B \text{ (in parallel):}\\
 \left\lfloor
\begin{array}{l}
\mu_y^{\text{pb}}(k+1) \rightarrow \text{ all prosumers } i \in \mc N_y^{\text{b}} \text{ on bus } y%
\end{array}%
\right.%
\end{array}%
\right.
\vspace*{.1em}
\end{array}
\right.%
\end{align*}
\smallskip
\hrule
\end{minipage}
}
\end{figure}%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%\centering
%\begin{subfigure}[b]{\columnwidth}
%%\begin{minipage}{\linewidth}
%\marginnote{\note{R1-6}}
%\edit{
%\hrule
%\smallskip
%\textsc{Algorithm 1}. Coordinated P2P Market Clearing
%\smallskip
%\hrule 
%\smallskip
%\text{Initialization}: For all prosumers $i \in \mc N$: set $\mu^{\text{tr}}_{(i,j)}(-1) = 0$, $\forall j \in \mc N_i$. DNO: set $\lambda^{\text{mg}}(0)\!=\!0$, $\mu^{\text{tg}}(0) \!=\! 0$, $\mu_y^{\text{pb}}(0)\! = \!0$, $\forall y \!\in\! \mc B$.
%
%\smallskip
% \text{Iterate until convergence $(k = 0,1,\ldots)$}
%\begin{align*}
%& \text{For all prosumers $i \in \mc I$ (in parallel):}\\
%\quad &\left|
%\begin{array}{l}
%\text{Local update via \textsc{Algorithm 2}:}\\
%\left\lfloor
%\begin{array}{l}
%\text{Set $\{\mu_{(i,j)}(k)\}_{j \in \mc N_i}$ as in \textsc{Alg. 2} (i)}\\
%\text{Set $u_i(k\!+\!1)$ as in \textsc{Alg. 2} (ii)}
%\end{array}
%\right.\\[1em]
% \text{Communication}\\
% \left\lfloor
%\begin{array}{l}
%\eta_i(k\!+\!1), \; p_i^{\text{mg}}(k\!+\!1) \rightarrow \text{ DNO}\\[.5em]
%\text{For all trading partners } j \in \mc N_i \text{ (in parallel):}\\
% \left\lfloor
%\begin{array}{l}
%p_{(i,j)}^{\text{tr}}(k\!+\!1) \rightarrow \text{ prosumer } j
%\end{array}
%\right.
%\end{array}
%\right.
%\end{array}
%\right.\\[.5em]
%%%%
%% DNO
%%%%
%& \text{DNO}\\
%\quad &\left|
%\begin{array}{l}
%\text{Central update via \textsc{Algorithm 3}:}\\
%\left\lfloor
%\begin{array}{l}
%\text{Set $u_{N+1}(k\!+\!1)$ as in \textsc{Alg. 3} (i)}\\
%\text{Set $\lambda^{\text{mg}}$, $\mu^{\text{tg}}$, $\{ \mu_y^{\text{pb}}(k\!+\!1)\}_{y \in\mc B}$ as in \textsc{Alg. 3} (ii)}
%\end{array}
%\right.\\[1em]
%\text{Communication (Gather \& Broadcast)}\\
% \left\lfloor
%\begin{array}{l}
%\sigma^{\mathrm{mg}},\, \lambda^{\text{mg}}, \mu^{\text{tg}}(k\!+\!1) \rightarrow \text{ all prosumers } i \in \mc N\\[.5em]
%\text{For all busses } y \in \mc B \text{ (in parallel):}\\
% \left\lfloor
%\begin{array}{l}
%\mu_y^{\text{pb}}(k+1) \rightarrow \text{ all prosumers } i \in \mc N_y^{\text{b}} \text{ on bus } y%
%\end{array}%
%\right.%
%\end{array}%
%\right.\\[1em]
%\end{array}
%\right.%
%\end{align*}
%\smallskip
%\hrule
%%\end{minipage}
%}
%\end{subfigure}
%\vspace*{2em}
%
%\begin{subfigure}[b]{0.37\textwidth}
%   \includegraphics[width=\linewidth]{figures/37bus_sys_f6}
%\end{subfigure}
%\caption{\edit{Information flow in Algorithm 1.}}
%\label{fig:IFC}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

Note that the cost functions in \eqref{eq:cost_def} are coupled only via the aggregative quantity $\sigma^{\textrm{mg}} = \sum_{i\in\mathcal{N}} p_{i}^{\mathrm{mg}}$ in  \eqref{eq:Aggr}, namely, the active load (i.e., the congestion) on the main grid. Therefore, for each agent $i \in \mc I$, we can define a function $\tilde J_i$ such that
\begin{equation}
J_i(u_i,  u_{-i})=\tilde J_i(u_i, \sigma^{\textrm{mg}}) .
\end{equation}
Games with such special structure are known as \textit{aggregative games} \cite{jensen2010aggregative}, and have received intense research interest, within the operations research and the automatic control communities \cite{paccagnan2019,belgioioso2020semi,belgioioso2020b,
gadjov2020single,bianchi2020fast}. 
\marginnote{\note{R1-5}}
\edit{
When the agents' cost functions depend linearly on the congestion (as for our P2P market model), v-GNEs are efficient. Specifically, the so-called \textit{price of anarchy}\cite{koutsoupias1999worst}, which quantifies how much selfish behaviour degrades the performance of a given system, tends to one (i.e., no performance degradation) as the agents population size grows \cite{paccagnan2018efficiency}.}

@@

\item a (dual) variable $\mu^{\text{tr}}_{(i,j)}$ for each trading partner $j \in \mc N_i$, whose function is to drive prosumer $i$'s and $j$'s power trade to agreement, i.e., the reciprocity constraints \eqref{eq:rep_const}, and can be interpreted from an economic perspective as a bilateral trading shadow price \cite[Section 2.4]{lecadre2020}.
\end{itemize}

\begin{figure}[t!]%
\begin{minipage}{\columnwidth}
\hrule
\smallskip
\textsc{Algorithm 1}. Distributed P2P Market Clearing
\smallskip
\hrule 
\smallskip
\text{Initialization}: For all prosumers $i \in \mc N$: set $\mu^{\text{tr}}_{(i,j)}(-1) = 0$, $\forall j \in \mc N_i$. DNO: set $\lambda^{\text{mg}}(0)\!=\!0$, $\mu^{\text{tg}}(0) \!=\! 0$, $\mu_y^{\text{pb}}(0)\! = \!0$, $\forall y \!\in\! \mc B$.

\smallskip
 \text{Iterate until convergence $(k = 0,1,\ldots)$}
\begin{align*}
& \text{For all prosumers $i \in \mc N$ (in parallel):}\\
\quad &\left\lfloor
\begin{array}{l}
\text{Local update via \textsc{Algorithm 2}:}\\
\left\lfloor
\begin{array}{l}
\text{Set $\{\mu_{(i,j)}(k)\}_{j \in \mc N_i}$ as in \textsc{Alg. 2} (i)}\\
\text{Set $u_i(k\!+\!1)$ as in \textsc{Alg. 2} (ii)}
\end{array}
\right.\\[1em]
 \text{Communication}\\
 \left\lfloor
\begin{array}{l}
\eta_i(k\!+\!1), \; p_i^{\text{mg}}(k\!+\!1) \rightarrow \text{ DNO}\\[.5em]
\text{For all trading partners } j \in \mc N_i \text{ (in parallel):}\\
 \left\lfloor
\begin{array}{l}
p_{(i,j)}^{\text{tr}}(k\!+\!1) \rightarrow \text{ prosumer } j
\end{array}
\right.
\end{array}
\right.
\vspace*{.1em}
\end{array}
\right.\\[.5em]
%%%
% DNO
%%%
& \text{DNO}\\
\quad &\left\lfloor
\begin{array}{l}
\text{Central update via \textsc{Algorithm 3}:}\\
\left\lfloor
\begin{array}{l}
\text{Set $u_{N+1}(k\!+\!1)$ as in \textsc{Alg. 3} (i)}\\
\text{Set $\lambda^{\text{mg}}$, $\mu^{\text{tg}}$, $\{ \mu_y^{\text{pb}}(k\!+\!1)\}_{y \in\mc B}$ as in \textsc{Alg. 3} (ii)}
\end{array}
\right.\\[1em]
\text{Communication (Gather \& Broadcast)}\\
 \left\lfloor
\begin{array}{l}
\sigma^{\mathrm{mg}},\, \lambda^{\text{mg}}, \mu^{\text{tg}}(k\!+\!1) \rightarrow \text{ all prosumers } i \in \mc N\\[.5em]
\text{For all busses } y \in \mc B \text{ (in parallel):}\\
 \left\lfloor
\begin{array}{l}
\mu_y^{\text{pb}}(k+1) \rightarrow \text{ all prosumers } i \in \mc N_y^{\text{b}} \text{ on bus } y%
\end{array}%
\right.%
\end{array}%
\right.
\vspace*{.1em}
\end{array}
\right.%
\end{align*}
\smallskip
\hrule
\end{minipage}

\vspace*{3em}

\centering
\includegraphics[width=.36\textwidth]{figures/37bus_sys_f7}
\smallskip
\caption{\edit{Information flow in Algorithm 1.}}
\label{fig:IFC}
\end{figure}%

In addition to the physical variables of the distribution network, i.e., $u_{N\!+\!1}$, the DNO stores in its local memory
\begin{itemize}
\item the (dual) variables $\lambda^{\text{mg}}$ and $\mu^{\text{tg}} $, that are associated with the main grid constraints \eqref{eq:p_mg_bound} and \eqref{eq:grid_const2}, respectively;

\item a (dual) variable $\mu_{y}^{\text{pb}}$ for each bus $y \in \mc B$, associated with the power balance constraint on bus $y$ \eqref{eq:p_bal_p}.
\end{itemize}
From an economic perspective, these variables can be interpreted as extra marginal losses imposed to the prosumers for the grid usage. From a control-theoretic perspective, they can be interpreted as states of discrete-time integrators driven by the violation of the network operational constraints \eqref{eq:Cconstr_def}.
}
%
%
%\begin{figure}[t]
%\marginnote{\note{R1-6}}
%\edit{
%\begin{minipage}{\columnwidth}
%\hrule
%\smallskip
%\textsc{Algorithm 1}. Coordinated P2P Market Clearing
%\smallskip
%\hrule 
%\smallskip
%\text{Initialization}: For all prosumers $i \in \mc N$: set $\mu^{\text{tr}}_{(i,j)}(-1) = 0$, $\forall j \in \mc N_i$. DNO: set $\lambda^{\text{mg}}(0)\!=\!0$, $\mu^{\text{tg}}(0) \!=\! 0$, $\mu_y^{\text{pb}}(0)\! = \!0$, $\forall y \!\in\! \mc B$.
%
%\smallskip
% \text{Iterate until convergence $(k = 0,1,\ldots)$}
%\begin{align*}
%& \text{For all prosumers $i \in \mc I$ (in parallel):}\\
%\quad &\left\lfloor
%\begin{array}{l}
%\text{Local update via \textsc{Algorithm 2}:}\\
%\left\lfloor
%\begin{array}{l}
%\text{Set $\{\mu_{(i,j)}(k)\}_{j \in \mc N_i}$ as in \textsc{Alg. 2} (i)}\\
%\text{Set $u_i(k\!+\!1)$ as in \textsc{Alg. 2} (ii)}
%\end{array}
%\right.\\[1em]
% \text{Communication}\\
% \left\lfloor
%\begin{array}{l}
%\eta_i(k\!+\!1), \; p_i^{\text{mg}}(k\!+\!1) \rightarrow \text{ DNO}\\[.5em]
%\text{For all trading partners } j \in \mc N_i \text{ (in parallel):}\\
% \left\lfloor
%\begin{array}{l}
%p_{(i,j)}^{\text{tr}}(k\!+\!1) \rightarrow \text{ prosumer } j
%\end{array}
%\right.
%\end{array}
%\right.
%\vspace*{.1em}
%\end{array}
%\right.\\[.5em]
%%%%
%% DNO
%%%%
%& \text{DNO}\\
%\quad &\left\lfloor
%\begin{array}{l}
%\text{Central update via \textsc{Algorithm 3}:}\\
%\left\lfloor
%\begin{array}{l}
%\text{Set $u_{N+1}(k\!+\!1)$ as in \textsc{Alg. 3} (i)}\\
%\text{Set $\lambda^{\text{mg}}$, $\mu^{\text{tg}}$, $\{ \mu_y^{\text{pb}}(k\!+\!1)\}_{y \in\mc B}$ as in \textsc{Alg. 3} (ii)}
%\end{array}
%\right.\\[1em]
%\text{Communication (Gather \& Broadcast)}\\
% \left\lfloor
%\begin{array}{l}
%\sigma^{\mathrm{mg}},\, \lambda^{\text{mg}}, \mu^{\text{tg}}(k\!+\!1) \rightarrow \text{ all prosumers } i \in \mc N\\[.5em]
%\text{For all busses } y \in \mc B \text{ (in parallel):}\\
% \left\lfloor
%\begin{array}{l}
%\mu_y^{\text{pb}}(k+1) \rightarrow \text{ all prosumers } i \in \mc N_y^{\text{b}} \text{ on bus } y%
%\end{array}%
%\right.%
%\end{array}%
%\right.
%\vspace*{.1em}
%\end{array}
%\right.%
%\end{align*}
%\smallskip
%\hrule
%\end{minipage}
%}
%\end{figure}%
